Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-02-22 ms.tex
editing

\subsection{Combined Heating Rate for Representative Star Formation History} 
\label{sec:combined}

The total gas heating rate, $\vwO$, includes contributions from
stellar winds, supernovae, pulsars, and SMBH feedback.  The strength
of each heating source will depend explicitely on the SMBH mass and
the star formation hstory in the galactic nuclear region, the latter
of which will also vary systematically with the galaxy mass and hence
with $\Mbh$.

Figure~\ref{fig:vwSourcesImp} shows the contributions of various
heating sources as a function of time $\tau_{\star}$ after a burst of
star formation for a black hole masses of $\Mbh = 10^{6} \Msun$ (top
panel) and $\Mbh = 10^{8} \Msun$ (bottom panel). The stellar wind
heating is calculated as described in appendix~\ref{app:windheat}. The
and Ia heating ($v_{w}^{\rm Ia}$) and compton heating ($v_{\rm
  w}^{\bullet}$) are calculated using equations~\eqref{eq:vIa}
and~\eqref{eq:vC} respectively. Both heating sources depend on the
value of the mass return parameter, $\eta$. We have plotted this for a
burst of star formation in Figure~\ref{fig:vwImp}.  The Ia heating rate
also depends be a convolution of the delay time distribution
(eq. [\ref{eq:DTD}]) and the star formation history. This convolution
reduces to the delay time distribution itself for a burst of star
formation. To account for the non-steady nature of the Ia heating we
multiply the Ia heating by $e^{-\rs/\rIa}$, where $\rIa$ is the radius
where the time between subsequent supernovae is equal to the local
dynamical time. The heating contribiution from
Ia supernovae will be suppressed if the time between successive
supernovae is much greater than the dynamical time at $\rs$. Note that
both $v_{\rm w}^{\bullet}$ and $v_{\rm w}^{\rm Ia}$ depend on the
total heating rate (all of the heating sources added in quadrature),
and so we have to solve a system of impllicit eqautions for these.
The MSP contribution is negligible for both cases and is not plotted.

The overall heating rate declines with time and after some time the
flow would become thermally unstable. The epoch when this occurs is
shown by dashed lines in Figure~\ref{fig:vwSourcesImp}. In calculating
this transition we have not included the compton heating. However,
this would not be able stabilize flow as it would not respond
immediately to cooling instabilities. Thus, if Compton
heating is the dominant source of heating we do not expect a steady,
thermally stable flow. Thermal instability is the norm in the burst
limit except for very early times ($\lsim 10^7$ years) and small
$\Mbh$ ($\Mbh\simeq 10^6 \Msun$).

>>>>

\subsection{Combined Heating Rate} 
\label{sec:combined}

The total gas heating rate, 
\be
\vwO = \sqrt{(v_{w}^{\star})^{2} + (v_{w}^{\rm MSP})^{2} + (v_{w}^{\rm Ia})^{2} + (v_{w}^{\bullet})^{2}},
\label{eq:vtot}
\ee
includes contributions from stellar winds, supernovae, pulsars, and radiative SMBH feedback (we neglect kinetic SMBH feedback for the moment, given its uncertain efficacy).  The strength of each heating source depends explicitly on the SMBH mass and the stellar population in the galactic nuclear region.  The latter could best be described by a single star burst formation in the past, or by a more continuous star formation history that itself varies systematically with the galaxy mass and hence $\Mbh$.

Figure~\ref{fig:vwSourcesImp} shows the contributions of various heating sources as a function of time $\tau_{\star}$ after a burst of star formation for black holes of mass $\Mbh = 10^{6} \Msun$ (top panel) and $\Mbh = 10^{8} \Msun$ (bottom panel). The heating and mass input due to stellar winds is calculated as described in Appendix~\ref{app:windheat} (Fig.~\ref{fig:vwImp}).  The heating from SN Ia ($v_{w}^{\rm Ia}$) and Compton heating ($v_{\rm w}^{\bullet}$) are calculated from equations~\eqref{eq:vIa} and~\eqref{eq:vC} respectively using the mass input parameter from stellar winds, $\eta$ (Fig.~\ref{fig:vwImp}).  SN Ia heating also depends on a convolution of the delay time distribution (DTD; eq. [\ref{eq:DTD}]) and the star formation history, which reduces to the DTD itself for a single star burst.  We crudely account for the expected reduction in Ia heating efficiency due to the non-steady nature of the SN Ia by attenuating the Ia heating rate by a factor $e^{-\rs/\rIa}$, that reduces to zero when the stagnation radius exceeds the radius, $r_{\rm Ia}$, at which the time between subsequent supernovae is equal to the local dynamical time (eq.~[\ref{eq:rIa}]).  Finally, note that $v_{\rm w}^{\bullet}$ and $v_{\rm w}^{\rm Ia}$ also depend on the total heating rate (eq.~[\ref{eq:vtot}]), requiring us to solve an implicit equation.  Heating from MSPs is found to be negligible compared to other sources of heating for fiducial parameters and hence is not plotted for clarity.

Figure~\ref{fig:vwSourcesImp} shows that for low mass black holes, stellar heating is the most important on timescales less than a few tens of millions of years.  Compton heating comes to dominate at later times due to the high stellar mass loss rates coupled with the higher accretion rates that accompany an overall decrease in the heating rate.  For higher mass black holes, Compton heating dominates at all epochs.  Because the total heating rate declines with time, after some point the flow becomes thermally unstable near the stagnation radius, and hence across all radii.  The epoch after this occurs according to the criterion $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}| \lesssim 1$ (eq.~[\ref{eq:cooling3}]) are shown as dashed lines in Figure~\ref{fig:vwSourcesImp}.  This transition is calculated not include Compton heating, since the latter cannot truly stabilize flow due to its inability to respond instantaneously to local changes in the thermal properties of the gas.  We thus see that thermally unstable flow is the norm in the single starburst case, except at very early times ($\lsim 10^7$ years) for small $\Mbh\simeq 10^6 \Msun$.
